\chapter{Testing and Discussion}

\section{Introduction}
A.U.R.A is a healthcare system for dementia patients requiring high reliability. It integrates React Native, FastAPI, MongoDB, FaceNet, Sarvam AI, Google Gemini, and Firebase. Testing ensures these layers function correctly to prevent poor patient experiences or missed SOS alerts.

\section{Testing Objectives}
\begin{itemize}
    \item \textbf{Unit Verification}: Ensure individual modules behave correctly.
    \item \textbf{Integration}: Validate communication between the app, backend, and external APIs.
    \item \textbf{Feature Functionality}: Confirm end-to-end performance of face recognition, SOS alerts, and medication management.
    \item \textbf{Performance \& Security}: Assess stability under concurrent use and enforce caregiver data restrictions.
    \item \textbf{Specialized Testing}: Verify Malayalam speech accuracy and SOS reliability.
\end{itemize}

\section{Testing Methods}
\begin{itemize}
    \item \textbf{Unit Testing}: Isolated testing of business logic using \textbf{pytest} and \textbf{Jest} with mock dependencies.
    \item \textbf{Functional Testing}: Manual and scripted flows validating features from the user perspective.
    \item \textbf{Integration Testing}: Verifying boundaries between components (e.g., App $\leftrightarrow$ Backend, Backend $\leftrightarrow$ Firebase).
    \item \textbf{Load Testing}: Using \textbf{Locust} to measure system stability and response times under concurrent usage.
\end{itemize}

\section{Test Cases}
\subsection{Unit Test Cases}
Tests confirmed face embedding generation, distress keyword flagging, medication formatting, and caregiver authentication (UT-01 to UT-10).

\subsection{Functional Test Cases}
Validated core workflows including Orito voice interactions, manual/voice SOS triggers, Memory Journal logging, and caregiver access controls (FT-01 to FT-12).

\subsection{Integration Test Cases}
Confirmed seamless data flow between the backend and external services like Google Gemini, Sarvam ASR/Bulbul, and Firebase (IT-01 to IT-10).

\subsection{Load Test Cases}
The system handled concurrent Orito messages, simultaneous SOS alerts, and ramp-up scenarios up to 50 users within acceptable performance thresholds (LT-01 to LT-08).

\section{Testing Summary}
All test cases passed, confirming A.U.R.A is stable and reliable for real-world deployment. Key findings include 100\% reliability in SOS delivery, effective caregiver privacy barriers, and stable backend performance under load. No critical defects were identified.

\chapter{Experimental Results}
This chapter presents the experimental results of the A.U.R.A. system through a series of screenshots from the patient, caregiver, and administrator application modules. It provides a detailed explanation of the functionalities within each interface, illustrating how the system facilitates interaction, monitoring, and management for all user roles.
\newpage
