
\chapter*{\centering Abstract}
\markboth{Custom Header}{Abstract}
\addcontentsline{toc}{chapter}{\numberline{}Abstract}%

Dementia is a progressive neurological disorder that causes severe memory loss, leaving
patients unable to recognize familiar faces, manage medication schedules, track daily
routines, or respond safely during emergencies — placing an enormous burden on both
patients and caregivers. Existing assistive solutions remain inaccessible to most due to high
cost, technical complexity, and lack of regional language support.
A.U.R.A (AI Assistive User Reminder Application) is an intelligent, affordable assistive
system developed to address these challenges. The system consists of the AURA Module,
a camera-equipped hardware unit currently prototyped using a computer webcam, which
performs real-time face detection and recognition using a custom-trained deep learning
model achieving above 99.66% accuracy, and a React Native mobile application serving as
the unified interface for patients and caregivers.
The application features Orito, a conversational AI assistant powered by the Google Gemini
API that communicates naturally in Malayalam. Orito enables patients to identify approach-
ing individuals by simply asking, query today’s medications and scheduled events, set
reminders through natural conversation, and automatically dispatch SOS alerts to registered
caregivers upon detecting patient distress. The backend is built using FastAPI (Python),
with all data — including facial embeddings, medication records, event schedules, care-
giver registrations, and SOS logs — stored in MongoDB for its flexible document-oriented
schema.
A.U.R.A implements a dual-role authentication model. Patients undergo an adaptive on-
boarding assessment on first login, allowing the application to optimize its interface and
interaction style based on their cognitive condition. Caregivers are authorized exclusively
via a Gmail OAuth mechanism pre-registered by the patient, with portal access restricted
solely to medication details and incoming SOS alerts to preserve patient privacy.
Built entirely on open-source frameworks and cloud AI services, A.U.R.A proves that
intelligent, culturally sensitive assistive technology can be developed affordably. Future
scope includes dedicated embedded AURA Module hardware, GPS-based patient tracking,
wearable fall detection, and expanded multilingual support, establishing A.U.R.A as a
scalable foundation for AI-driven dementia care.

 \setcounter{equation}{0}
\chapter{Introduction}
 Dementia is one of the most challenging neurodegenerative conditions affecting millions of elderly individuals worldwide. It progressively deteriorates memory, cognitive function, and the ability to perform routine daily tasks. Among those affected, patients in regional communities face an even greater challenge — most available assistive technologies are designed for English-speaking, technically literate users, leaving a significant gap for people who communicate in regional languages such as Malayalam. Families caring for dementia patients in Kerala are forced to maintain near-constant physical presence, as existing solutions fail to address the full spectrum of daily needs these patients face.
The rapid advancement of artificial intelligence, mobile computing, and speech processing technologies has opened new possibilities for building intelligent assistive systems that are affordable, accessible, and culturally appropriate. However, these technologies have largely not been brought together in a way that directly serves the dementia care context, particularly for regional language users in India.
A.U.R.A — AI Assistive User Reminder Application — is developed in response to this gap. It is a smart, integrated system that combines real-time face recognition, a Malayalam-speaking conversational AI assistant named Orito, medication and event management, a Memory Journal, automatic distress detection, bidirectional location tracking, and caregiver alerting — all within a single mobile application built on React Native, backed by a FastAPI server and MongoDB database, and powered by a custom fine-tuned FaceNet model, Google Gemini API, and Sarvam AI's speech services.
  \section{ Problem Statement}
  Dementia patients often struggle with cognitive decline, making it difficult to recognize family members and adhere to medication schedules. This is further compounded by a significant language barrier, as most existing assistive interfaces are English-only, rendering them inaccessible to elderly users in Kerala who are more comfortable in their native language. Beyond these challenges, safety risks remain a serious concern, as dementia-related wandering and the inability to use smartphones during emergencies frequently lead to delayed responses. Adding to this, caregivers and families lack reliable tools for remote monitoring, and the high cost of commercial care platforms places an additional burden on households already managing the demands of dementia care.
  


  \section{Objectives and Scope}
  A.U.R.A aims to provide an affordable, voice-driven assistant named Orito, built around four core areas. For Malayalam AI, the system leverages Google Gemini and Sarvam AI to enable fully natural speech-in and speech-out interactions in Malayalam. In terms of recognition and memory, A.U.R.A employs a fine-tuned FaceNet model for real-time face identification, alongside a Memory Journal that logs all conversations for the patient to review later. On the safety and logistics front, the system includes automated distress detection that requires no button input, bidirectional GPS tracking, and tools for managing medications and events. Finally, secure access is handled through Firebase Google Sign-In, which grants caregivers restricted access to safety-related data while keeping personal journals and face data strictly private.
 
 
\section{Methodology Overview}
The system utilizes a React Native mobile app and a FastAPI backend integrated with MongoDB. At the core of its AI pipeline, it integrates Sarvam ASR and TTS models with Google Gemini to provide contextual and natural dialogue in Malayalam. Complementing the software is the AURA Module, a local hardware component such as a Raspberry Pi or a laptop, which handles real-time face recognition. For critical situations, an emergency protocol featuring a hybrid keyword and LLM system reliably detects distress and instantly triggers Firebase push notifications to caregivers. Overall, this architecture intelligently combines open-source tools and free-tier cloud services to ensure the entire solution remains highly affordable and accessible for the average Indian family.
  