\chapter{System Study}
This chapter presents a detailed study of the technologies, software components, and hardware elements that form the foundation of A.U.R.A. Each technology was selected based on its suitability for the specific requirements of the system --- real-time face recognition, Malayalam speech processing, context-aware conversational AI, and low-cost deployment for dementia patients in Kerala.

\section{Technologies}

\subsection{React Native (TypeScript) --- Cross-Platform Mobile Application Framework}
React Native is an open-source framework developed by Meta that enables building native mobile applications for both Android and iOS from a single codebase using JavaScript or TypeScript. Unlike hybrid web-based approaches, React Native renders actual native UI components, giving the application the performance and look and feel of a platform-native app. TypeScript adds static type checking on top of JavaScript, reducing runtime errors and improving code maintainability across a large multi-feature codebase.

A.U.R.A's mobile application is built entirely in React Native with TypeScript. The same application serves both the patient and caregiver under different login roles --- the patient interface with Orito, the Memory Journal, medication reminders, face recognition feed, and SOS button, and the caregiver interface with location tracking, SOS alerts, and patient health reports. Building both roles in a single React Native codebase avoids maintaining two separate applications and ensures consistent behavior across Android and iOS devices. Expo was used alongside React Native to streamline the development, testing, and build process.

\subsection{FastAPI (Python) --- Backend Server Framework}
FastAPI is a modern, high-performance Python web framework for building APIs, designed around Python type hints and based on the ASGI standard. It is one of the fastest Python frameworks available, with performance comparable to Node.js and Go. FastAPI automatically generates interactive API documentation, supports asynchronous request handling natively, and integrates cleanly with Python's machine learning and AI ecosystem --- making it a natural choice for a backend that coordinates multiple AI services simultaneously.

The A.U.R.A backend is built entirely on FastAPI running under Uvicorn as the ASGI server. Every request from the mobile app --- patient messages to Orito, face registration, medication management, event scheduling, SOS triggering, location updates, and caregiver data retrieval --- is handled by the FastAPI server. The server orchestrates calls to Google Gemini, Sarvam ASR, Sarvam Bulbul, MongoDB, and Firebase in a single coordinated pipeline. FastAPI's asynchronous capabilities ensure that multiple concurrent requests from the patient and caregiver apps are handled without blocking.

\subsection{MongoDB --- NoSQL Document Database}
MongoDB is a document-oriented NoSQL database that stores data as flexible JSON-like documents rather than rigid relational tables. This schema flexibility makes it well suited to applications where different data types have different structures and where requirements evolve over time. MongoDB supports efficient querying, indexing, and embedding of related data within documents, making it practical for applications that mix structured data (medication schedules, event reminders) with unstructured data (conversation logs, embedding vectors).

A.U.R.A uses MongoDB as its primary database. The following collections are maintained: patient profiles and credentials, FaceNet embedding vectors (stored instead of raw images for privacy), medication records, event schedules, Memory Journal entries (timestamped conversation history between the patient and Orito), caregiver registrations, AURA Module recognition results, real-time location data, and SOS alert logs. MongoDB's document model allows each of these varied data structures to coexist naturally without schema conflicts, and its flexibility accommodates the ongoing evolution of the system's features.

\subsection{Google Gemini API --- Large Language Model for Conversational AI}
The Google Gemini API provides access to Google's family of large Transformer-based language models capable of understanding and generating natural language across multiple languages. Gemini supports multimodal input, long context windows, and is specifically designed for integration into applications via API. Its in-context learning capability --- whereby the model's behavior can be directed entirely through the content of the prompt without any fine-tuning --- is central to A.U.R.A's design.

Gemini powers Orito, the Malayalam-speaking conversational AI assistant at the heart of A.U.R.A. Before every patient interaction, the backend fetches the patient's complete current context from MongoDB --- today's medication schedule, upcoming events, daily routines, and the latest face recognition result from the AURA Module --- and injects all of this into the Gemini prompt alongside the patient's message. This ensures that every response Orito gives is specific to that patient's actual real-time situation rather than a generic answer. Gemini also simultaneously performs semantic distress detection on every message, enabling automatic SOS alerting when genuine distress is identified.

\subsection{Sarvam ASR --- Automatic Speech Recognition for Indian Languages}
Sarvam ASR is a Transformer-based automatic speech recognition system developed by Sarvam AI, specifically designed and trained for Indian languages including Malayalam. It handles the linguistic diversity of Indian speech --- regional accents, code-mixed sentences, noisy real-world environments, and elderly speech patterns --- far more accurately than general-purpose ASR systems trained on western language data. The system follows the same self-supervised Transformer-based ASR principles established by wav2vec 2.0, adapted specifically for the Indian language domain.

Sarvam ASR is the speech-to-text engine for Orito. When the patient speaks to Orito in Malayalam, the audio captured by the mobile app is sent to Sarvam ASR, which transcribes it accurately into Malayalam text. This transcribed text is then sent to the FastAPI backend as the patient's message for processing by Gemini. Since every patient interaction begins with this transcription step, the accuracy and robustness of Sarvam ASR directly determines the overall usability of the system for elderly Malayalam-speaking dementia patients.

\subsection{Sarvam Bulbul --- Neural Text-to-Speech for Indian Languages}
Sarvam Bulbul is a neural text-to-speech model developed by Sarvam AI for Indian languages, emphasising naturalness, prosodic expressiveness, and handling of code-mixed content such as English medication names embedded within Malayalam sentences. Its pipeline normalises and phonemises input text, generates a mel-spectrogram through a neural acoustic model, and converts it to audible speech through a neural vocoder --- producing speech that sounds significantly more natural and warm than legacy concatenative TTS systems.

Bulbul is the voice output engine for Orito. Every Malayalam response generated by Gemini is passed to Bulbul for synthesis into natural-sounding speech, which is then returned to the mobile app and played back to the patient. This voice output capability is indispensable for dementia patients who struggle to read text on a screen. Bulbul's naturalness and warmth make interactions with Orito feel genuinely conversational rather than robotic, which is critical for user acceptance among elderly patients.

\subsection{Firebase Authentication and Cloud Messaging}
Firebase is Google's mobile and web application development platform. Firebase Authentication provides a secure, ready-built identity management system supporting multiple sign-in methods including Google Sign-In. Firebase Cloud Messaging (FCM) is a cross-platform push notification service that delivers messages reliably to both Android and iOS devices even when the receiving app is in the background.

A.U.R.A uses Firebase for two distinct purposes. First, caregiver login is handled through Firebase Authentication with Google Sign-In. The patient pre-registers a caregiver's Gmail address in their profile; when the caregiver attempts to log in, Firebase authenticates their Google account and the backend verifies that their Gmail matches a registered caregiver entry in MongoDB --- ensuring only explicitly pre-approved contacts can access patient data. Second, Firebase Cloud Messaging delivers SOS push notifications to all registered caregivers' devices the moment an SOS is triggered, whether by voice distress detection or the manual button, ensuring immediate alert delivery regardless of whether the caregiver's app is open.

\subsection{OpenCV --- Computer Vision for Face Detection}
OpenCV (Open Source Computer Vision Library) is an open-source computer vision and machine learning library widely used for real-time image and video processing. It provides efficient implementations of face detection algorithms, frame capture from cameras, image preprocessing operations, and video stream management. OpenCV runs efficiently on a wide range of hardware including resource-constrained devices like Raspberry Pi.

In A.U.R.A's AURA Module, OpenCV is responsible for the first two stages of the face recognition pipeline. It continuously reads frames from the camera and applies face detection to identify whether a face is present in the frame and where it is located. The detected face region is cropped and preprocessed by OpenCV before being passed to the FaceNet model for embedding generation. OpenCV's efficient real-time processing is what enables the AURA Module to run the full recognition pipeline locally on the device without streaming raw video over the network.

\subsection{FaceNet --- Deep Metric Learning for Face Recognition}
FaceNet is a deep neural network developed by Schroff et al. that maps face images to a compact 128-dimensional embedding space using deep metric learning. It is trained with triplet loss --- using anchor, positive, and negative image triples --- to ensure that embeddings of the same person cluster together while embeddings of different people are pushed apart. Once trained, recognition requires only a fast distance comparison between embedding vectors, with no retraining needed to add new identities. A custom fine-tuned version of FaceNet, adapted on a locally representative dataset for Kerala, is used in A.U.R.A.

FaceNet is the core face recognition algorithm of the AURA Module. When OpenCV crops a detected face from the camera feed, FaceNet generates a 128-dimensional embedding for that face. This embedding is compared against all stored embeddings in MongoDB using Euclidean distance and cosine similarity. If the closest match falls within a defined threshold, the person is identified by name and relationship; otherwise they are flagged as unknown. Only the final result --- not any raw image or video --- is sent to the backend. All registered persons are stored as embedding vectors only, making the system both storage-efficient and privacy-preserving.

\section{Software Components}

\subsection{Mobile Application}
The A.U.R.A mobile application is a single React Native (TypeScript) app built with Expo that serves both patients and caregivers under separate login roles. The patient-facing interface includes the Orito voice assistant screen, the AURA Module Connect section for discovering and pairing with the face recognition module, a camera preview showing the live feed from the AURA Module, the My People face registration gallery, the Memory Bank (Memory Journal) for browsing all past Orito conversations, a Reminders and Tasks section, a Calendar view for scheduled events and medications, and a Settings panel. The caregiver-facing interface includes a Care Center dashboard showing linked patients, a Live Location map powered by OpenStreetMap tiles, a Medications desk for reviewing and editing patient treatment plans, an SOS History log, Activity Reports with mood and journal signals, and a notifications management panel.

The two roles are cleanly separated at the login stage --- patients authenticate with a custom username and password system, while caregivers authenticate via Firebase Google Sign-In. The backend enforces role-based access on every API call, ensuring caregivers can only access medications, reminders, SOS history, and location data, and have no access to the Memory Journal or any other private patient information.

\subsection{Backend Server}
The FastAPI backend is the central nervous system of A.U.R.A. It exposes a RESTful API that the mobile app and AURA Module connect to over the network. Its core responsibilities are as follows. For patient interactions, it receives transcribed text from the app, fetches the patient's full context from MongoDB, sends the combined prompt to Gemini, performs distress detection on the response, saves the exchange to the Memory Journal, synthesises speech via Bulbul, and returns audio to the app. For face recognition, it receives the identification result from the AURA Module and stores it in MongoDB for use in the next Orito interaction. It also handles face registration (generating and storing FaceNet embeddings), medication and event CRUD operations, caregiver registration and verification, real-time location updates from both patient and caregiver devices, and SOS alert dispatch via Firebase Cloud Messaging.

\subsection{Orito --- Conversational AI Assistant}
Orito is the Malayalam-speaking AI assistant that serves as the primary human-computer interface for the patient. It is not a standalone model but a configured interaction layer built on top of the Gemini API, Sarvam ASR, and Sarvam Bulbul. Its behaviour is shaped entirely by the system prompt and context injection performed by the FastAPI backend before every call. Orito is configured to always respond in Malayalam, to address the patient by name, to use the patient's current medication and event data in its answers, to incorporate the latest AURA Module face recognition result when the patient asks about a person, to extract structured information from natural conversation using LLM-assisted Named Entity Recognition for saving reminders, and to detect distress semantically and trigger SOS alerts automatically. Every Orito interaction is saved to the Memory Journal automatically.

\subsection{Memory Bank (Memory Journal)}
The Memory Bank is a dedicated section of the patient app that stores and displays the complete chronological history of every conversation between the patient and Orito. Each entry records the patient's spoken message, Orito's response, and a timestamp. The Memory Bank also stores face identification events and journal summaries generated by Orito. Patients can search their memories by people or date, and filter by category --- People, Journal, or AI Insights. The Memory Bank is private to the patient and is completely inaccessible to caregivers.

\subsection{SOS Alert System}
A.U.R.A implements a dual-mode SOS system. The first mode is voice-triggered: every message the patient sends to Orito is simultaneously evaluated by Gemini using a two-layer hybrid approach --- fast keyword detection flags potential distress words first, and Gemini's semantic understanding then validates whether the message constitutes a genuine emergency, reducing false positives while ensuring genuine distress is never missed. The second mode is a large manual SOS button prominently visible on the patient's home screen, for situations where the patient cannot speak. When either mode triggers, the backend immediately queries MongoDB for all registered caregivers linked to that patient and dispatches push notifications to their devices via Firebase Cloud Messaging. SOS events are logged with trigger type, timestamp, GPS location, and caregiver notification status.

\section{Hardware}

\subsection{AURA Module --- Current Prototype (Laptop-Based)}
In the current prototype, the AURA Module is implemented on a laptop. The laptop runs the face recognition software locally using its built-in webcam as camera input. The full recognition pipeline --- OpenCV frame capture, face detection and cropping, FaceNet embedding generation, and embedding comparison against MongoDB --- executes locally on the laptop. Only the final recognition result (the matched person's name and relationship, or ``unknown'') is sent to the FastAPI backend over the local network via HTTP. No raw video or image data is transmitted over the network at any point.

The mobile app connects to the AURA Module through a simple discovery and tap-to-connect flow --- the app scans for available AURA Modules on the local network and the user taps the detected device to pair, similar to connecting a Bluetooth device. Once connected, the app displays a live camera preview feed from the module so the patient or family member can verify the camera angle without any technical knowledge. This laptop-based prototype demonstrates the complete system behavior and pipeline with no compromise to the core logic.

\subsection{AURA Module --- Future Hardware (Raspberry Pi-Based Standalone Device)}
The production version of the AURA Module will be a purpose-built, self-contained standalone hardware device with the following specifications. The processing unit will be a Raspberry Pi, which has sufficient computational capacity to run OpenCV and the fine-tuned FaceNet model locally without any cloud dependency. The camera will be a dedicated Raspberry Pi camera module for continuous face capture. Power will be supplied by a 10,000 mAh rechargeable battery pack, making the device fully wireless and portable --- the patient's family can place it anywhere in the home without needing a nearby power outlet. Connectivity will use the same wireless local network discovery and pairing flow already implemented in the current prototype.

The future device requires no setup beyond powering it on and tapping to connect from the app. The entire processing pipeline remains local to the device, preserving the same privacy-by-design architecture of the current prototype. The estimated hardware cost of this future device keeps the total system accessible to ordinary families --- consistent with A.U.R.A's design goal of being a genuinely affordable alternative to expensive commercial dementia care solutions.

\subsection{Patient Smartphone}
The patient's Android or iOS smartphone runs the A.U.R.A mobile application. The smartphone's microphone captures the patient's spoken Malayalam for Sarvam ASR transcription, and its speaker plays back Orito's synthesised voice responses via Sarvam Bulbul. The phone's GPS module provides real-time location data that the backend stores in MongoDB and serves to the caregiver's location tracking view. The smartphone does not perform any local AI inference --- all recognition, language, and speech processing is handled by the AURA Module or the backend --- keeping the app lightweight and accessible on low-specification devices.
