\begin{thebibliography}{999}
\addcontentsline{toc}{chapter}{\numberline{}References}
\raggedright

\bibitem{bib1}
World Health Organization,
 “Dementia,” Sep. 2023. Accessed: Dec. 29, 2025.
 [Online]. Available: https://www.who.int/news-room/fact-sheets/detail/dementia


\bibitem{bib2}
Google,
 “Gemini API Documentation.”
 [Online]. Available: https://ai.google.dev/

\bibitem{bib3}
SOS Alert System,
 “Emergency SOS: How it works.”
 [Online]. Available: https://support.google.com/android/answer/9319337

\bibitem{bib4}
CareYaya’s MedaCareLLM system,
 “AI glasses support dementia patients,” Nov. 05, 2024.
 [Online]. Available: https://healthcare-digital.com/technology-and-ai/careyayas-medacarellm-ai-glasses-support-dementia-patients

\bibitem{bib5}
MongoDB,
 “MongoDB Documentation.”
 [Online]. Available: https://www.mongodb.com/



%  \bibitem{bib10}
% ”KSEB Bill Calculator”, [Online], Available:
% https://bills.kseb.in/
% \bibitem{bib3}
%  ”React Native”, [Online], Available:
%  https://reactnative.dev/
%   \bibitem{bib5}
% ”IoT Smart Electricity Energy Meter with ESP32 and Blynk 2.0”, [Online], Available: https://how2electronics.com/iot-smart-electricity-energy-meter-with-esp32-blynk-2-0/
\bibitem{bib6}
F. Schroff, D. Kalenichenko, and J. Philbin,
 ``FaceNet: A unified embedding for face recognition and clustering,''
 in Proc. IEEE Conf. Computer Vision and Pattern Recognition (CVPR), 2015, pp. 815--823.Available: https://arxiv.org/abs/1503.03832

\bibitem{bib7}
A. Baevski, Y. Zhou, A. Mohamed, and M. Auli,
 ``wav2vec 2.0: A framework for self-supervised learning of speech representations,''
 in Advances in Neural Information Processing Systems (NeurIPS), vol. 33, 2020.Available: https://arxiv.org/abs/2006.11477

\bibitem{bib8}
Sarvam AI,
 ``Bulbul: Text-to-Speech for Indian Languages,'' 2024.
 [Online]. Available: https://www.sarvam.ai

\bibitem{bib9}
T. B. Brown et al.,
 ``Language models are few-shot learners,''
 in Advances in Neural Information Processing Systems (NeurIPS), vol. 33, 2020.Available:https://arxiv.org/abs/2005.14165

\bibitem{bib10}
Firebase,
 ``Firebase Authentication and Cloud Messaging Documentation.''
 [Online]. Available: https://firebase.google.com/docs

\bibitem{bib11}
Meta,
 ``React Native Documentation.''
 [Online]. Available: https://reactnative.dev/docs/getting-started

\bibitem{bib12}
FastAPI,
 ``FastAPI Documentation.''
 [Online]. Available: https://fastapi.tiangolo.com/

\bibitem{bib13}
OpenCV,
 ``OpenCV Documentation.''
 [Online]. Available: https://docs.opencv.org/

\end{thebibliography}